---
title: "Improving the Accuracy of Aggregate Growth Measures: Statistical Methods and Practical Challenges"
subtitle: "Discussion of Presentations"
author: "Damian Betebenner"
institute: "Center for Assessment"
date: "April 17th, 2022 (updated: `r toOrdinal::toOrdinalDate(Sys.Date())`)"
output:
  xaringan::moon_reader:
    lib_dir: libs
    nature:
      highlightStyle: github
      highlightLines: true
      countIncrementalSlides: false
      titleSlideClass: ["center", "middle", "my-title"]

---

```{r setup, include=FALSE}
options(htmltools.dir.version = FALSE)
```

```{r share-again, echo=FALSE}
xaringanExtra::use_share_again()
```

```{r style-share-again, echo=FALSE}
xaringanExtra::style_share_again(
  share_buttons = c("twitter"),
  background = "#009C49",
  foreground = "#FFFFFF"
)
```

```{r xaringan-logo, echo=FALSE}
xaringanExtra::use_logo(
  image_url = "https://centerforassessment.github.io/assets/favicon.png",
  width = "55px",
  height = "64px",
  position = xaringanExtra::css_position(top = "0.5em", right = "0.5em"),
  link_url = "https://www.nciea.org"
)
```

```{r xaringan-tile-view, echo=FALSE}
xaringanExtra::use_tile_view()
```

```{r xaringanExtra-clipboard, echo=FALSE}
xaringanExtra::use_clipboard()
```

# Empirical Best Linear Prediction

--

- An excellent set of presentations/paper.

--

- A major contribution to the way in which states and other organizations should consider aggregating individual growth scores.

--

- I commend the authors for creating a GitHub based R package ([schoolgrowth](https://github.com/jrlockwood/schoolgrowth)) with the non-trivial analytics.

```{r echo=FALSE}
magick::image_read_pdf("Figures/github_repo_schoolgrowth.pdf")
```

---

# Presentation 1:

## Improving Accuracy and Stability of Aggregate Student Growth Measures Using Empirical Best Linear Prediction

--

- Results clearly show the superior of EBLP over the simple average except in a few circumstances (e.g. some large districts).

--

- Huge benefits in terms of accuracy and precistion for small groups (e.g., schools)

--

- I found the fact that the approach does not involve shrinkage but improves accuracy to be somewhat counter-intuitive.

--

- A feature of the EBLP approach that "makes it work" are the "patterns" described in the paper. Could you elaborate on exactly how patterns are defined.

---

# Presentation 2:

## Applications of the Empirical Best Linear Prediction to Aggregate Growth Measures in Two States

--

- Excellent stress test of the approach using real data

--

- Comparison showing increasing stability including additional subjects and lag scores was compelling

--

- The Decision Points that were laid out at the end of the presentation are well stated and highly practical

--

- To what extent does pooling results across years influence stability versus EBLP?

---

# Presentation 3:

## Adopting a Growth Measure in California

--

- Great to see California moving forward to having student growth. Better late than never.

--

- Their journey is illustrative of the numerous pitfalls that can accompany assessment and accountability initiatives.

--

- I'm curious. You mention in 2018 during your review that the growth model did not "Provide the answers/information that our various educational partners wanted". What answers/information did the partners want.



---

# Overall Comments

--

- Growth aggregation methods:

--

	- Started with median

--

	- Percentage of observations above some threshold

--

	- Mean

--

	- EBLP


- IMHO there's a inverse relationship between the technical soundness of the aggregation approach and ease of communcating to non-technical audiences.

--

- Data as flashlight instead of a hammer (accountability)

--

- Would this technique be amenable to any data. For example student scaled scores? (as a substitute for percent proficient)
